---
title: "Data 211 Project"
author: "Kofi Owusu"
date: "2026-04-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load library
library(ggplot2)
library(readxl)
```


```{r}
#data
sleep_data <- read_excel("sleep_data.xlsx", range = "A1:B22")
```

## Summary Statistics
```{r}
summary(sleep_data$sleep_hours)
sd(sleep_data$sleep_hours)
```
## Visualization

```{r}
# Line Graph
ggplot(sleep_data, aes(x = day, y = sleep_hours)) +
  geom_line(color = "purple", size = 1) +
  geom_point(color = "black", size = 2.5) +
  geom_hline(yintercept = mean(sleep_data$sleep_hours),
             linetype = "dashed", color = "red") +
  labs(
    title = "Daily Sleep Hours – April 5 to April 25",
    x = "Day",
    y = "Hours of Sleep"
  ) +
  scale_x_continuous(breaks = 1:21) +
  theme_minimal()
```
The line graph shows my nightly sleep hours fluctuating throughout the 21-day period. There is no consistent pattern. Some nights I slept well above average while others dropped significantly. Most nights fall between 6 and 8 hours, with day 9 being the highest at 10.5 hours and day 8 being the lowest at 3.5 hours


```{r}
#BoxPlot
ggplot(sleep_data, aes(y = sleep_hours)) +
  geom_boxplot(fill = "purple", color = "black", width = 0.4) +
  geom_hline(yintercept = 8, linetype = "dashed", color = "red") +
  labs(
    title = "Sleep Hours: Spread and Variability",
    x = " ",
    y = "Hours of Sleep"
  ) +
  theme_minimal() 
```
The box plot shows that the median sleep is 7.0 hours and the middle 50% of nights fall roughly between 5.5 and 8.0 hours. The dashed line at 8 hours shows that my median sits below the recommended amount. The spread of the box indicates moderate variability in my sleep habits.


```{r}
#Scatterplot
ggplot(sleep_data, aes(x = day, y = sleep_hours)) +
  geom_point(color = "purple", size = 3) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  geom_hline(yintercept = 8, linetype = "dashed", color = "purple") +
  labs(
    title = "Sleep Trend Over 21 Days",
    x = "Day",
    y = "Hours of Sleep"
  ) +
  scale_x_continuous(breaks = 1:21) +
  theme_minimal()
```
The trend line shows a slight upward slope, suggesting my sleep hours marginally increased over the 21-day period. However, the high variability among the data points indicates this trend is weak and not conclusive.


## Statistical Test

```{r}
t.test(sleep_data$sleep_hours, mu = 8, alternative = "two.sided")

```
## Interpretation
The one-sample t-test produced a t-statistic of -2.902 and a p-value of 0.0088, which is below the significance level of 0.05. I reject the null hypothesis. There is strong statistical evidence that my average nightly sleep of 6.88 hours is significantly less than the recommended 8 hours.
